Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2021-10-13 methods_paper/fabl-paper.tex
minor updates.

	Record linkage is the task of identifying duplicate records across multiple data sources, often in the absence of a unique identifier \citep{christen_2012}. This is an increasingly important task in ``data cleaning,'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others. In this paper, we consider bipartite record linkage, which is merging two databases that contain duplication across the two databases, but do not contain any duplications within each database.

	%Much of the statistical literature focuses on bipartite record linkage \citep{fellegi_theory_1969, jaro1989, Winkler1988, belin_1995, larsen_2001, liseo_2011,  herzog2007data, gutman_bayesian_2013, sadinle_bayesian_2017}.

	Many statistical record linkage methods are extensions of the seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}. Specifically, Fellegi and Sunter created comparison vectors for each pair of records in the data and independently classified those pairs as a match or a non-match using a likelihood ratio test. Recent work in the statistical literature has extended this approach for a wide variety of applications \citep{winkler1991application, fair2004generalized, wagner2014person, gill2003english}. Although the need to compute comparisons for all record pairs initially restricted this family of methods, recent work by \cite{enamorado2019using} used innovative hashing techniques to scale the approach to handle large datasets.

@@

	Alternatively, one can embed one-to-one matching requirements into the model specification itself, at an additional computational cost. \cite{larsen2005} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space. \cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method was so computationally intensive it could only be used to link databases of less than 100 records. Most recently, \cite{sadinle_bayesian_2017} proposed \emph{Beta Record Linkage} (\texttt{BRL}), using an innovative prior over the space of bipartite matchings to strictly enforce one-to-one reqiurements throughout his Gibbs sampler. Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeller can weight the relative importance of false positives and false negatives, and identify records pairings to be be decided through clerical review. Although it was shown to work on larger tasks than previous one-to-one methods, \texttt{BRL} becomes slow when working with larger datasets, and unusable when the number of pairwise record comparisons becomes too large.

	In this paper, we propose \emph{fast beta linkage} (\texttt{fabl}), which extends the the Beta Record Linkage model for increased efficiency and scalability. Specifically, we relax the one-to-one matching requirement of \texttt{BRL} and propose independent priors over the matching space, creating a ``many-to-one'' model for record linkage. This allows us to (1) employ hashing techniques that hasten calculations and reduce computational costs; (2) compute the pairwise record comparisons over large datasets via parallel computing; and (3) reduce memory costs through our proposed method of \emph{storage efficient indexing}. Throughout, we argue that even in cases where a bipartite matching is desired, our proposed approach (with a simple post-processing procedure) provides equivalently accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed. Open source software to use \texttt{fabl} in \texttt{R} is available through Github.

	In what follows, Section \ref{sec:review-of_prior-work} reviews more thoroughly the Fellegi Sunter and Beta Record Linkage methods. Then, Section \ref{sec:fast-beta-linkage} provides the model specification under \texttt{fabl} and derives the Gibbs sampler for posterior inference. Next, Section \ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of the linkage tasks we can undertake. Lastly, Sections \ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and a case study of homicides from the El Salvadoran Civil War. Section \ref{discussion} concludes our paper with a discussion of directions for future work.

@@

The simplest way to compare two records is to check for agreement or disagreement, and this is commonly used for categorical variables. For more complex measurements, we can take into account partial agreement to more richly characterize the comparison; for numerical data, we can use absoluate difference, and for text data, we can use string distance metrics such as Levenstein or Jaro-Winkler distance. We can then propose thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}. Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f}$. Following convention, $I_{f0}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement). Thus, we can construct comparison vectors in the following way: 
$$\gamma_{ij}^f = \ell \; \text{if} \; \mathcal{S}_f(i,j) \in I_{f\ell}.$$

The choice of $I_{f\ell}$ are application specific, which we discuss in our simulation and case study.

@@

Importantly, this independence means that we have relaxed the one-to-one requirement from \texttt{BRL}; our sampler does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\). This behavior may be desirable in its own right; \cite{Newcombe1962b} were interested in matching birth records to marriage certificates, a situation in which such ``many-to-one'' matchings are reasonable. In situations when a bipartite estimate is desired, we can use a simple post processing procedure shown in \ref{bayes-estimate} to resolve illegal matchings.

@@

We seek to minimize the posterior expected loss, given by 
$$\mathbb{E}[L(\hat{\bm{Z}}, \bm{Z})|\Gamma^{obs}] = \sum_{\bm{Z}} L(\hat{Z_j}, Z_j)P(\bm{Z}|\Gamma^{obs})$$. 
In this paper, we adopt losses $\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2$, inducing the intuitive decision rule

@@

To achieve a Bayes estimate that fulfills one-to-one matching requirement, we simply minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$. In the event that we have two records $j$ and $j'$ such that both $P(Z_j = i |\Gamma) > \frac{1}{2}$ and $ P(Z_{j'} = i |\Gamma) > \frac{1}{2}$, this constraint means we accept the match with the highest posterior probability, and declare the other to have no match.  A similar approach can be see in the most probable maximal matching sets used by \citep{steorts_2013} to match records to latent entities, and is justified within this Bayesian framework.

>>>>

	Record linkage is the task of identifying duplicate records across multiple data sources, often in the absence of a unique identifier \citep{christen_2012}. This is an increasingly important task in ``data cleaning,'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others. \textcolor{red}{In this paper, we consider bipartite record linkage, which merges two databases together that contain duplications across but not within the respective databases \citep{fellegi_theory_1969, jaro1989, Winkler1988, belin_1995, larsen_2001, liseo_2011,  herzog2007data, gutman_bayesian_2013, sadinle_bayesian_2017}.}
	%Much of the statistical literature focuses on bipartite record linkage \citep{fellegi_theory_1969, jaro1989, Winkler1988, belin_1995, larsen_2001, liseo_2011,  herzog2007data, gutman_bayesian_2013, sadinle_bayesian_2017}. 
	Many statistical record linkage methods are extensions of the seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}. Specifically, Fellegi and Sunter created comparison vectors for each pair of records in the data and independently classified those pairs as a match or a non-match using a likelihood ratio test. Recent work in the statistical literature has extended this approach for a wide variety of applications \citep{winkler1991application, fair2004generalized, wagner2014person, gill2003english}. Although the need to compute comparisons for all record pairs initially restricted this family of methods, recent work by \cite{enamorado2019using} used innovative hashing techniques to scale the approach to handle large datasets.

@@

	Alternatively, one can embed one-to-one matching requirements into the model specification itself, at an additional computational cost. \cite{larsen2005} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space. \cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method was so computationally intensive it could only be used to link databases of less than 100 records. Most recently, \cite{sadinle_bayesian_2017} proposed the \emph{Beta Record Linkage} model (\texttt{BRL}), using an innovative prior over the space of bipartite matchings to strictly enforce one-to-one reqiurements throughout his Gibbs sampler. Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeller can weight the relative importance of false positives and false negatives, and identify records pairings to be be decided through clerical review. Although it was shown to work on larger tasks than previous one-to-one methods, \texttt{BRL} becomes slow when working with larger datasets, and unusable when the number of pairwise record comparisons becomes too large.

	In this paper, we propose \emph{fast beta linkage} (\texttt{fabl}), which extends the the Beta Record Linkage model for increased efficiency and scalability. Specifically, we relax the one-to-one matching requirement of \texttt{BRL} and propose independent priors over the matching space, creating a ``many-to-one'' model for record linkage. This allows us to (1) employ hashing techniques that hasten calculations and reduce computational costs; (2) compute the pairwise record comparisons over large datasets via parallel computing; and (3) reduce memory costs through our proposed method of \emph{storage efficient indexing}. Throughout, we argue that even in cases where a bipartite matching is desired, our proposed approach (with a simple post-processing procedure) provides equivalently accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed. Open source software to use \texttt{fabl} in \texttt{R} is available through Github at \textcolor{red}{BK: Add the URL even if it's private at the moment: \url{}}.

	\textcolor{red}{The remainder of the paper is as follows.} Section~\ref{sec:review-of_prior-work} reviews \textcolor{red}{the prior work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}.} In section~\ref{sec:fast-beta-linkage}, we propose our the \texttt{fabl} model specification. \textcolor{red}{Also, we derive the corresponding conditional distributions and provide a Gibbs sampler for posterior inference.}  Section \ref{bayes-estimate} \textcolor{red}{Brian: add in what you do in this section.}
	Section~\ref{sec:efficiency} \textcolor{red}{proposes our approach} for hashing and  and storage efficient indexing used to increase the speed of calculations and the scale of the linkage tasks we can undertake. In addition, we prove the computational complexity of \texttt{fabl}. Sections~\ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and a case study of homicides from the El Salvadoran Civil War. Section~\ref{discussion} concludes our paper with a discussion of directions for future work.

@@

The simplest way to compare two records is to check for agreement or disagreement, and this is commonly used for categorical variables. For more complex measurements, we can take into account partial agreement to more richly characterize the comparison; for numerical data, we can use absoluate difference, and for text data, we can use string distance metrics such as Levenstein or Jaro-Winkler distance. We can then propose thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}. Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f}$. Following convention, $I_{f0}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement). Thus, we can construct comparison vectors in the following way: 
$\gamma_{ij}^f = \ell \; \text{if} \; \mathcal{S}_f(i,j) \in I_{f\ell}.$
The choice of $I_{f\ell}$ are application specific, which we discuss in our simulation and case study.

@@

Importantly, this independence means that we have relaxed the one-to-one requirement from \texttt{BRL}; our sampler does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\). This behavior may be desirable in its own right; \cite{Newcombe1962b} were interested in matching birth records to marriage certificates, a situation in which such ``many-to-one'' matchings are reasonable. In situations when a bipartite estimate is desired, we can use a simple post processing procedure shown in  section~\ref{bayes-estimate} to resolve illegal matchings.

@@

We seek to minimize the posterior expected loss, given by 
$$\mathbb{E}[L(\hat{\bm{Z}}, \bm{Z})|\Gamma^{obs}] = \sum_{\bm{Z}} L(\hat{Z_j}, Z_j)P(\bm{Z}|\Gamma^{obs}).$$
In this paper, we adopt losses $\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2$, inducing the intuitive decision rule

@@

To achieve a Bayes estimate that fulfills one-to-one matching requirement, we simply minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$. In the event that we have two records $j$ and $j'$ such that both $P(Z_j = i |\Gamma) > \frac{1}{2}$ and $ P(Z_{j'} = i |\Gamma) > \frac{1}{2}$, this constraint means we accept the match with the highest posterior probability, and declare the other to have no match.  A similar approach can be see in the most probable maximal matching sets used by \cite{steorts_bayesian_2016} to match records to latent entities, and is justified within this Bayesian framework.
